Minimax-optimal nonparametric regression with graph Laplacians

### 2020-09-16 notes/graph_regression2.tex
Completed an edit of the first draph of our graph regression paper; time to merge.

In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Rd$, and 
\begin{equation}
\label{eqn:random_design_regression}
Y_i = f_0(X_i) + \varepsilon_i
\end{equation}
with independent Gaussian noise $\varepsilon_i \sim N(0,1)$. Our goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether or not $f_0 = 0$, i.e whether or not there is a signal present.

In graph based nonparametric regression, the idea is to perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$. For a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$. The weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ is an $n \times n$ matrix with entries
\begin{equation*}
\label{eqn:neighborhood_graph}
{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\Biggr),
\end{equation*}
and the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$.  Our star operator, the neighborhood graph Laplacian, can then be written as
\begin{equation}
\label{eqn:graph_Laplacian}
\Lap_{n,r} = \bf{D} - \bf{W}
\end{equation}
The two graph-based regression methods we will consider are \emph{Laplacian eigenmaps} and \emph{Laplacian smoothing}. Each method uses $\Lap_{n,r}$ in a different manner to construct an estimator of $f_0$. Letting $v_1(G_{n,r}),\ldots,v_n(G_{n,r})$ denote eigenvectors of $\Lap_{n,r}$, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ \citep{belkin2003} is given by 
\begin{equation}
\label{eqn:laplacian_eigenmaps}
\wh{f}_{\textrm{LE}} = \sum_{k = 1}^{\kappa} \Dotp{Y}{v_k(G_{n,r})}_n v_{k}(G_{n,r})
\end{equation}
where $\kappa \in \{1,...,n\}$ is a tuning parameter. (Here $\dotp{u}{v}_n = \frac{1}{n}\sum_{i = 1}^{n} u_i v_i$ for vectors $u,v \in \Reals^n$.) The Laplacian smoothing estimator $\wt{f}_{\LS}$ \citep{smola2003} is a penalized least squares estimator, given by
\begin{equation}
\label{eqn:laplacian_smoothing}
\wt{f}_{\LS}= \min_{f \in \Reals^n} \Bigl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_{n,r}^s f \Bigr\}
\end{equation}
where $\rho > 0$ acts a tuning parameter $f^T \Lap_{{n,r}}^s f$. Then, assuming~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are reasonable estimators of $f_0$, the squared empirical norms
\begin{equation}
T_{\LE} = \Bigl\|\wh{f}_{\LE}\Bigr\|_n^2 \label{eqn:laplacian_eigenmaps_test}
\end{equation}
and
\begin{equation}
T_{\LS} = \Bigl\|\wt{f}_{\LS}\Bigr\|_n^2 \label{eqn:laplacian_smoothing_test}
\end{equation}
are in turn reasonable statistics to test whether or not $f_0 = 0$.

There already exist a vast number of methods for both the nonparametric testing and, especially, the nonparametric estimation problems. In fact, as we will see in Section~\ref{sec:problem_setup_and_background} some of these methods are actually quite analogous to the graph Laplacian estimation and testing methods we have just introduced. Nevertheless, methods such as~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} have a deserved place in the nonparametric regression toolbox. Graph-based methods such as these have many advantages: they are easy to implement, fast and stable to compute, generalizable to non-standard data---for instance, text or images, or really any data modality on which it is possible to define a kernel--- and non-standard learning paradigms---such as the semi-supervised setting---and last but not least, have enjoyed widespread practical success.

For these reasons, a substantial body of work has arisen assessing the statistical properties of graph Laplacian based methods. Taken as a whole, this work views the graph Laplacian matrix as an estimator of a weighted continuum Laplace operator $\Delta_P$ (see equations~\eqref{eqn:laplace_operator} and~\eqref{eqn:laplace_operator_weighted} in Section~\ref{sec:problem_setup_and_background}), and establishes that it is in fact a \emph{consistent} estimator of $\Delta$, in various respects. For instance,  \citet{belkin03,belkin05} show that when $\Pbb$ is the uniform distribution over $\Xset$, $\Lap_{n,r}f \to \Delta f$ for sufficiently smooth functions $f$, \citet{lafon04,hein05} extend the statement to hold when $\Pbb$ is non-uniform, \citet{singer06} sharpens the convergence rates, and \cite{gine06} establish a CLT. On the other hand, \citet{belkin07} prove that when $\Pbb$ is the uniform distribution over $\mathcal{X}$, the eigenvalues and eigenvectors of $\Lap_{n,r}$ converge to those of $\Delta$ when $r_n \to 0$ as $n \to \infty$, \citet{garciatrillos18} extend this result to hold for general $\Pbb$, and \cite{calder2019} obtain sharper rates at which $r_n$ may converge to $0$.

These developments notwithstanding, the following fundamental theoretical question remains unanswered: are graph-based methods for nonparametric regression not merely consistent but in fact \emph{optimal}, in the minimax sense? In this paper, we confirm that they are. We show that when the regression function $f_0$ is suitably smooth---by which we mean that it belongs to a first-order Sobolev ball; roughly speaking, that it possesses weak derivative $Df$ bounded in $L^2(\Xset)$ norm---both~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are with high probability asymptotically rate-optimal estimators, and that both~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} are asymptotically rate-optimal test statistics. We also show that under appropriate conditions, these estimators and test statistics can leverage additional assumed structure, either on the regression function $f_0$ or distribution $P$. Our fundamental message is a reassuring one: the practitioner who wishes to use neighborhood graph procedures, for any of the various reasons outline above, may do so without fear that they are sacrificing statistical optimality in the process.

\begin{itemize}
	\item \textbf{Motivate graph Laplacians.}

	\item \textbf{Literature review.} 
	\item \textbf{Organization.}
\end{itemize}

\paragraph{Notation.}
\begin{itemize}
	\item We write $\mathbb{Z}^d = \mathbb{Z} \otimes \cdots \otimes \mathbb{Z}$ (the tensor product is taken a total of $d$ times); the sequence space $\ell^2(\mathbb{Z}^d)$ contains those vectors $\theta: \mathbb{Z}^d \to \Reals$ for which $\sum_{k \in \mathbb{Z}^d} \theta_k^2 < \infty$. 
	\item We denote by $\Leb^2(\Xset)$ the set of square integrable functions $f:\Xset \to \Reals$. \textcolor{red}{Holder}.
	\item For functions $f,g$ which map from $\Xset$ to $\Reals$, we write $\dotp{f}{g}_n := \frac{1}{n}\sum_{i = 1}^{n} f(X_i) g(X_i)$, which we refer to as the \emph{empirical inner product}, and $\norm{f}_n^2 = \dotp{f}{f}_n$, which we refer to as the (squared) \emph{empirical norm}. For vectors $v \in \Reals^n$, we write $\norm{v}_p^p := \sum_{i = 1}^{n} \abs{v_i}^p$.
	\item For a graph $G$, we use $\Lap_G$ to denote the Laplacian of $G$. The matrix $\Lap_G$ is a square, symmetric, positive semi-definite matrix, with $n$ eigenvalues $\lambda_k(G)$ and eigenvector $v_k(G)$ pairs, defined according to
	\begin{equation*}
	\Lap_G v_k(G) = \lambda_k(G) v_k(G), ~~\frac{1}{n} \norm{v_k(G)}_2^2 = 1 
	\end{equation*}
	and ordered so that $0 \leq \lambda_1(G) \leq \ldots \leq \lambda_n(G)$. For convenience, when dealing with the neighborhood graph $G_{n,r}$ we denote $\Lap_{n,r} = \Lap_{G_{n,r}}$. 
	\item When convenient we write $\mathbf{X} = (X_1,\ldots,X_n)$. 
\end{itemize}

@@

Before we get to our main results, we will review some of the relevant literature on minimax optimal nonparametric estimation and testing of Sobolev functions. We will also briefly recall some (of the many) estimators and tests which achieve these rates: in particular the \emph{Fourier series projection} and \emph{smoothing spline} methods. As we will see, Laplacian eigenmaps and Laplacian smoothing can be seen as graph-based, i.e. discrete, counterparts to these ``continuum'' methods.

We start by giving a (very) brief exposition of Sobolev spaces, with an emphasis on the material relevant in the statistical context.

@@

Roughly speaking, Sobolev spaces contain functions $f \in \Leb^p(\Xset)$ with derivatives which themselves belong to $\Leb^p(\Xset)$. More formally, for given integers $s$ and $p > 0$, the Sobolev space $W^{s,p}(\Xset)$ consists of all functions $f \in \Leb^p(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^p(\Xset)$. The Sobolev-$\{s,p\}$ seminorm $\seminorm{f}_{W^{s,p}(\Xset)}$ and norm $\norm{f}_{W^{s,p(\Xset)}}$ are then given by 
\begin{equation*}
\seminorm{f}_{W^{s,p}(\Xset)}^p = \sum_{\abs{\alpha} = s}\int_{\mathcal{X}} \Bigl|\bigl(D^{\alpha}f\bigr)(x)\Bigr|^p \,dx, ~~ \norm{f}_{W^{s,p}(\Xset)}^p = \sum_{k = 0}^{s} \seminorm{f}_{W^{k,p}(\Xset)}^p
\end{equation*}
and for a given $M > 0$, the corresponding ball is $W^{s,p}(\Xset, M) = \set{f: \norm{f}_{W^{s,p}(\Xset)} \leq M}$. In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space, and we adopt the usual convention of writing $H^s(\Xset) = W^{s,p}(\Xset)$ and $H^s(\Xset,M) = W^{s,2}(\Xset,M)$; we will confine our attention to this case hereafter. One special fact worth mentioning: when $\Xset$ is sufficiently smooth, for all Sobolev spaces $H^s(\Xset)$ the Holder space $C^{\infty}(\Xset)$ is dense in $H^s(\Xset)$. Said differently, for each integer $s$, the Sobolev space $H^s(\Xset)$ is simply the completion of $C^{\infty}(\Xset)$ in the norm $\norm{\cdot}_{H^s(\Xset)}$.

Periodic functions $f \in H^s(\Xset)$ admit a representation in terms of Fourier coefficients. Define the \emph{Sobolev ellipsoids} $\Theta_s(M) \subseteq \ell^2(\mathbb{Z}^d)$ to be
\begin{equation*}
\Theta_s(M) = \Bigl\{\theta \in \ell_2(\mathbb{Z}^d): \sum_{k \in \mathbb{Z}^d} \theta_k^2 \abs{k}^{2s} \leq M^2\Bigr\}.
\end{equation*}
To see the connection between $H^s(M)$ and $\Theta_s(M)$, for the moment let us fix the domain $\Xset = [0,1]^d$, and write $\set{\phi_k}_{k \in \mathbb{Z}^d}$ for the tensor product trigonometric polynomial basis of $\Leb^2(\Xset)$. Then the periodic Sobolev balls
\begin{equation*}
H_{\textrm{per}}^s(\Xset,M) = \Bigl\{\sum_{k \in \mathbb{Z}^d} \theta_k \phi_k: \theta_k \in \Theta_s(M) \Bigr\}
\end{equation*}
satisfy $H_{\textrm{per}}^s(\Xset,M) \subseteq H^s(\Xset,M')$ for a sufficiently large value of $M_1 > 0$. On the other hand, any $1$-periodic function $f \in H^s(\Xset,M)$ also belongs to $H_{\textrm{per}}^s(\Xset,M_0)$ for a sufficiently large value of $M_0 > 0$.

@@

\textcolor{red}{(TODO 1)}: Refer also to the Sobolev embedding theorem, for at least three reasons: (1) to facilitate later discussion of why smoothing splines are ill-posed when $2s < d$; (2) to make explicit why the statistical problem isn't ill-posed when $2s < d$ in the random design setting; (3) to make clear why the testing problem is hopeless when $4s < d$.

\subsection{Nonparametric regression: Fourier series and smoothing splines.}
\label{subsec:continuum_methods}
When the regression function $f_0$ is assumed to belong to a Sobolev space, two approaches particularly well-suited for nonparametric regression are orthogonal projection onto the Fourier basis, and smoothing splines. In the former, we estimate the Fourier coefficients $\wh{\theta}_k := \dotp{Y}{\phi_k}_n$, with the resultant function estimate $\wh{f}_{\OS}$ defined as
\begin{equation}
\label{eqn:orthogonal_series}
\wh{f}_{\OS} = \sum_{k: \abs{k} \leq \kappa} \wh{\theta}_k \phi_k
\end{equation}
for some $\kappa \geq 0$. In contrast, the smoothing spline estimator $\wt{f}_{\SM}$ is a variational estimator, defined as the solution to the following optimization problem,
\begin{equation}
\label{eqn:smoothing_spline}
\wt{f}_{\SM} = \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \seminorm{f}_{H^{s}(\Xset)}^2\biggr\}
\end{equation}
with the minimum taken over all $f \in H^s(\Xset)$.

The orthogonal series estimator $\wh{f}_{\OS}$ is closely connected with the Laplacian eigenmaps estimator $\wh{f}_{\LE}$. Just as $v_k(G_{n,r})$ are eigenvectors of the graph Laplacian $\Lap_{n,r}$, so too are $\phi_k$ eigenfunctions of a particular differential operator $\Delta: C^2(\Xset) \to \Leb^2(\Xset)$, known as the \emph{Laplace} operator and formally defined as
\begin{equation}
\label{eqn:laplace_operator}
\Delta g := -\mathrm{div}(\nabla g),
\end{equation}
where $\mathrm{div}$ stands for the divergence, and $\nabla$ for the gradient. As mentioned previously, when the radius $r \to 0$ as $n \to \infty$ at an appropriate rate, the graph Laplacian $\Lap_{n,r}$ in fact converges to a weighted Laplace operator $\Delta_P$, given by
\begin{equation}
\label{eqn:laplace_operator_weighted}
\Delta_{P} g := -\frac{1}{p}\mathrm{div}(p^2 \nabla g),
\end{equation}
with the eigenvectors $v_k(G_{n,r})$ converging in an appropriate sense to eigenfunctions of $\Delta_{P}$. The upshot is that when $P$ is the uniform measure---and thus the eigenfunctions of $\Delta_P = \Delta$ are in fact the trigonometric basis functions $\phi_k$---for an appropriate sequence of radii $r(n) \to 0$ and any fixed $\kappa \geq 0$ the estimator $\wh{f}_{\OS}$ is in fact the limit of $\wh{f}_{\LE}$, in the sense that
\begin{equation}
\label{eqn:laplacian_eigenmaps_consistency}
\abs{\wh{f}_{\LE}(X_i) - \wh{f}_{\OS}(X_i)} \overset{P}{\to} 0~~\textrm{as $n \to \infty$.}
\end{equation} 
for each $i \in \mathbb{N}$.

>>>>

In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Rd$, and 
\begin{equation}
\label{eqn:random_design_regression}
Y_i = f_0(X_i) + \varepsilon_i
\end{equation}
with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise. Our goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether $f_0 = 0$, i.e whether there is any signal present.

In graph-based nonparametric regression, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$. The neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$. The $n \times n$ weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ encodes proximity between pairs of samples; for a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the entries $\mathbf{W}_{ij}$ are given by
\begin{equation*}
\label{eqn:neighborhood_graph}
{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\Biggr).
\end{equation*}
Then the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$, and our star operator, the neighborhood graph Laplacian, can be written as
\begin{equation}
\label{eqn:graph_Laplacian}
\Lap_{n,r} = \bf{D} - \bf{W}
\end{equation}
The two graph-based regression methods we will consider are \emph{Laplacian eigenmaps} and \emph{Laplacian smoothing}. Each method uses $\Lap_{n,r}$ in a different manner to construct an estimator of $f_0$. Letting $v_1(G_{n,r}),\ldots,v_n(G_{n,r})$ denote eigenvectors of $\Lap_{n,r}$, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ \citep{belkin2003} is given by 
\begin{equation}
\label{eqn:laplacian_eigenmaps}
\wh{f}_{\textrm{LE}} = \sum_{k = 1}^{\kappa} \biggl\{\Bigl(\frac{1}{n}\sum_{i = 1}^{n} Y_i v_{k,i}(G_{n,r})\Bigr) v_{k}(G_{n,r})\biggr\}
\end{equation}
where $\kappa \in \{1,...,n\}$ is a tuning parameter. The Laplacian smoothing estimator $\wt{f}_{\LS}$ \citep{smola2003} is a penalized least squares estimator, given by
\begin{equation}
\label{eqn:laplacian_smoothing}
\wt{f}_{\LS}= \min_{f \in \Reals^n} \biggl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_{n,r}^s f \biggr\}
\end{equation}
where $\rho > 0$ acts a tuning parameter on the penalty $f^T \Lap_{{n,r}}^s f$. Assuming~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are reasonable estimators of $f_0$, the squared empirical norms
\begin{equation}
T_{\LE} = \frac{1}{n}\Bigl\|\wh{f}_{\LE}\Bigr\|_2^2 \label{eqn:laplacian_eigenmaps_test}
\end{equation}
and
\begin{equation}
T_{\LS} = \frac{1}{n}\Bigl\|\wt{f}_{\LS}\Bigr\|_2^2 \label{eqn:laplacian_smoothing_test}
\end{equation}
are in turn reasonable statistics to test whether or not $f_0 = 0$.

Graph-based methods have a deserved place in the nonparametric regression toolbox. Of course, there exist a considerable range of methods for both the nonparametric testing and, especially, the nonparametric estimation problems. In fact, as we will see in Section~\ref{sec:problem_setup_and_background} some of these more classical methods are actually quite analogous to~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing}. Nevertheless, graph-based methods have many strengths: they are easy to implement; fast and stable to compute; generalizable to non-standard data---for instance, text or images, or really any data modality on which it is possible to define a kernel--- and non-standard learning paradigms---such as the semi-supervised setting. Last but not least, they have enjoyed widespread practical success.

For these reasons, a substantial body of work has arisen assessing the statistical properties of graph-based methods. Taken as a whole, this work views the graph Laplacian matrix as an estimator of a weighted continuum Laplace operator $\Delta_P$ (see equations~\eqref{eqn:laplace_operator} and~\eqref{eqn:laplace_operator_weighted} in Section~\ref{sec:problem_setup_and_background}), and establishes that when $n \to \infty$ and $r(n) \to 0$, it is in fact a \emph{consistent} estimator of $\Delta_P$. This notion of consistency is meant in a few different respects. For instance,  \citet{belkin03,belkin05} show that when $P$ is the uniform distribution over $\Xset$, $\Lap_{n,r}f \to \Delta f$ for sufficiently smooth functions $f$, \citet{lafon04,hein05} extend the statement to hold when $P$ is non-uniform, \citet{singer06} sharpens the convergence rates, and \cite{gine06} prove a central limit theorem. In another line of work, \citet{belkin07} prove that when $P$ is the uniform distribution over $\mathcal{X}$, the eigenvalues and eigenvectors of $\Lap_{n,r}$ converge to those of $\Delta_P$, \citet{garciatrillos18} extend this result to hold for when $P$, and \cite{calder2019} sharpen the rates at which $r(n)$ may converge to $0$.

These developments notwithstanding, the following fundamental theoretical question remains unanswered: are graph-based methods for nonparametric regression not merely consistent but in fact \emph{optimal}, in the minimax sense? In this paper, we confirm that they are. We establish that when the regression function $f_0$ is suitably smooth---by which we mean that it belongs to a first-order Sobolev ball; roughly speaking, that it possesses weak derivative $Df$ bounded in $L^2(\Xset)$ norm---both~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are with high probability asymptotically rate-optimal estimators, and that both~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} are asymptotically rate-optimal test statistics. We also show that these estimators and test statistics can leverage additional assumed structure, either on the regression function $f_0$ or the distribution $P$. Our fundamental message is a reassuring one: the practitioner who wishes to use neighborhood graph procedures, for any of the various reasons outlined above, may do so without fear that they are sacrificing statistical optimality in the process.

\textcolor{red}{(TODO 1)}: Outline organization of paper here.

\paragraph{Notation.}
\begin{itemize}
	\item We write $\mathbb{Z}^d = \mathbb{Z} \otimes \cdots \otimes \mathbb{Z}$---with the tensor product being taken a total of $d$ times--- and likewise $\mathbb{N}^d = \mathbb{N} \otimes \cdots \otimes \mathbb{N}$. The sequence space $\ell^2(\mathbb{Z}^d)$ then contains those vectors $\theta: \mathbb{Z}^d \to \Reals$ for which $\sum_{k \in \mathbb{Z}^d} \theta_k^2 < \infty$. 
	\item For $1 \leq p < \infty$, we denote by $\Leb^p(\Xset)$ the set of functions $f:\Xset \to \Reals$ with bounded $p$th moment $\int_{\Xset} \bigl(f(x)\bigr)^p \,dx < \infty$. When $p = \infty$, we let $\Leb^{\infty}(\Xset)$ contain those functions $f$ for which $\sup_{x \in \Xset} |f(x)| < \infty$.
	\item For functions $f,g: \Xset \to \Reals$, we write $\dotp{f}{g}_n := \frac{1}{n}\sum_{i = 1}^{n} f(X_i) g(X_i)$, which we refer to as the \emph{empirical inner product}, and $\norm{f}_n^2 = \dotp{f}{f}_n$, which we refer to as the (squared) \emph{empirical norm}. For vectors $v \in \Reals^n$, we write $\norm{v}_p^p := \sum_{i = 1}^{n} \abs{v_i}^p$. When there is no chance of being confused, we will sometimes associate a function $f$ with its evaluations $(f(X_1),\ldots,f(X_n)) \in \Reals^n$.
	\item In general, for a graph $G$ we use $\Lap_G$ to denote the Laplacian of $G$. The matrix $\Lap_G$ is a square, symmetric, positive semi-definite matrix, with $n$ eigenvalues $\lambda_k(G)$ and eigenvector $v_k(G)$ pairs, defined according to
	\begin{equation*}
	\Lap_G v_k(G) = \lambda_k(G) v_k(G), ~~\frac{1}{n} \norm{v_k(G)}_2^2 = 1 
	\end{equation*}
	and ordered so that $0 \leq \lambda_1(G) \leq \ldots \leq \lambda_n(G)$. As we've already seen, to reduce notational clutter we write $\Lap_{n,r}$ instead of $\Lap_{G_{n,r}}$. 
	\item For sequences $a_n, b_n$, we write $a_n \lesssim b_n$ when there exists a constant $C > 0$ such that $a_n \leq C b_n$ for all $n$, and $a_n \asymp b_n$ when $a_n \lesssim b_n$ and $b_n \lesssim a_n$. 
	\item When convenient we write $\mathbf{X} = (X_1,\ldots,X_n)$. 
\end{itemize}

@@

Before we get to our main results, we will review some of the relevant literature on minimax optimal estimation and testing of Sobolev functions. We will also recall some (of the many) estimators and tests which achieve these rates: in particular the \emph{Fourier series projection} and \emph{smoothing spline} methods. As we will see, one can interpret Laplacian eigenmaps and Laplacian smoothing as graph-based, i.e. discrete, counterparts to these ``continuum'' methods.

@@

We start by briefly introducing Sobolev spaces, emphasizing only the material relevant to our statistical context. Roughly speaking, Sobolev spaces contain functions $f \in \Leb^p(\Xset)$ with derivatives which themselves belong to $\Leb^p(\Xset)$. More formally, for given integers $s$ and $p > 0$, the Sobolev space $W^{s,p}(\Xset)$ consists of all functions $f \in \Leb^p(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ exists and belongs to $\mathcal{L}^p(\Xset)$. For such functions, the $(s,p)$-Sobolev seminorm $\seminorm{f}_{W^{s,p}(\Xset)}$ and norm $\norm{f}_{W^{s,p(\Xset)}}$ are given by 
\begin{equation*}
\seminorm{f}_{W^{s,p}(\Xset)}^p = \sum_{\abs{\alpha} = s}\int_{\mathcal{X}} \Bigl|\bigl(D^{\alpha}f\bigr)(x)\Bigr|^p \,dx, ~~ \norm{f}_{W^{s,p}(\Xset)}^p = \sum_{k = 0}^{s} \seminorm{f}_{W^{k,p}(\Xset)}^p
\end{equation*}
and for a given $M > 0$, the $(s,p)$-Sobolev ball is $W^{s,p}(\Xset, M) = \set{f: \norm{f}_{W^{s,p}(\Xset)} \leq M}$. In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space, and we adopt the usual convention of writing $H^s(\Xset) = W^{s,p}(\Xset)$ and $H^s(\Xset,M) = W^{s,2}(\Xset,M)$; we will confine our attention to this case hereafter.

Periodic functions $f \in H^s(\Xset)$ admit a representation in terms of Fourier coefficients. Define the \emph{Sobolev ellipsoids} $\Theta_s(M) \subseteq \ell^2(\mathbb{Z}^d)$ to be
\begin{equation*}
\Theta_s(M) = \Bigl\{\theta \in \ell_2(\mathbb{Z}^d): \sum_{k \in \mathbb{Z}^d} \theta_k^2 \abs{k}^{2s} \leq M^2\Bigr\}.
\end{equation*}
To see the connection between $H^s(M)$ and $\Theta_s(M)$, for the moment let us fix the domain $\Xset = [0,1]^d$, and write $\set{\phi_k}_{k \in \mathbb{Z}^d}$ for the trigonometric basis of $\Leb^2(\Xset)$. Then the periodic Sobolev balls
\begin{equation*}
H_{\textrm{per}}^s(\Xset,M) = \Bigl\{\sum_{k \in \mathbb{Z}^d} \theta_k \phi_k: \theta_k \in \Theta_s(M) \Bigr\}
\end{equation*}
satisfy $H_{\textrm{per}}^s(\Xset,M) \subseteq H^s(\Xset,M_1)$ for a sufficiently large value of $M_1 > 0$. On the other hand, any $1$-periodic function $f \in H^s(\Xset,M)$ also belongs to $H_{\textrm{per}}^s(\Xset,M_0)$ for a sufficiently large value of $M_0 > 0$.

@@

\textcolor{red}{(TODO 1)}: Introduce the Sobolev embedding theorem, for at least three reasons: (1) to facilitate later discussion of why smoothing splines are ill-posed when $2s < d$; (2) to make explicit why the statistical problem isn't ill-posed when $2s < d$ in the random design setting; (3) to make clear why the testing problem is hopeless when $4s < d$. Conclude with one special fact worth mentioning: regardless of $s$ and $d$, when $\Xset$ is sufficiently smooth for any Sobolev spaces $H^s(\Xset)$ the set of infinitely differentiable functions $C^{\infty}(\Xset)$ is dense in $H^s(\Xset)$. Said differently, for each integer $s$, the Sobolev space $H^s(\Xset)$ is simply the completion of $C^{\infty}(\Xset)$ in the norm $\norm{\cdot}_{H^s(\Xset)}$.

\subsection{Nonparametric regression: Fourier series and smoothing splines}
\label{subsec:continuum_methods}
Orthogonal projection onto the Fourier basis and smoothing splines are two approaches particularly well-suited for nonparametric regression over Sobolev spaces. In the former approach, we estimate the Fourier coefficients $\wh{\theta}_k := \dotp{Y}{\phi_k}_n$, with the resulting function estimate
\begin{equation}
\label{eqn:orthogonal_series}
\wh{f}_{\OS} = \sum_{k: \abs{k} \leq \kappa} \wh{\theta}_k \phi_k
\end{equation}
for some $\kappa \geq 0$. In contrast, the smoothing spline estimator $\wt{f}_{\SM}$ is a variational estimator, defined as the solution to the following optimization problem,
\begin{equation}
\label{eqn:smoothing_spline}
\wt{f}_{\SM} = \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \seminorm{f}_{H^{s}(\Xset)}^2\biggr\}
\end{equation}
with the minimum taken over all $f \in H^s(\Xset)$.

The orthogonal series estimator $\wh{f}_{\OS}$ is closely connected with the Laplacian eigenmaps estimator $\wh{f}_{\LE}$. Just as $v_k(G_{n,r})$ are eigenvectors of the graph Laplacian $\Lap_{n,r}$, so too are $\phi_k$ eigenfunctions of a particular differential operator $\Delta: C^2(\Xset) \to \Leb^2(\Xset)$, known as the \emph{Laplace} operator and formally defined as
\begin{equation}
\label{eqn:laplace_operator}
\Delta g := -\mathrm{div}(\nabla g),
\end{equation}
where $\mathrm{div}(\cdot)$ stands for the divergence, and $\nabla$ for the gradient. As mentioned previously, when the radius $r \to 0$ as $n \to \infty$ at an appropriate rate, the graph Laplacian $\Lap_{n,r}$ in fact converges to a weighted Laplace operator $\Delta_P$, given by
\begin{equation}
\label{eqn:laplace_operator_weighted}
\Delta_{P} g := -\frac{1}{p}\mathrm{div}(p^2 \nabla g);
\end{equation}
likewise the eigenvectors $v_k(G_{n,r})$ converge to eigenfunctions of $\Delta_{P}$. The upshot is that when $P$ is the uniform measure on $\Xset = [0,1]^d$---and thus the eigenfunctions of $\Delta_P = \Delta$ are precisely the trigonometric basis functions $\phi_k$--- and $\kappa \geq 0$ is fixed, the estimator $\wh{f}_{\OS}$ is in fact the limit of $\wh{f}_{\LE}$, in the sense that
\begin{equation}
\label{eqn:laplacian_eigenmaps_consistency}
\abs{\wh{f}_{\LE}(X_i) - \wh{f}_{\OS}(X_i)} \overset{P}{\to} 0~~\textrm{as $n \to \infty$ and $r(n) \to 0$,}
\end{equation} 
for each $i \in \mathbb{N}$.

By viewing $\Delta_P$ as the continuum limit of~$\Lap_{n,r}$, one can also tie the penalty in the smoothing spline estimator $\wh{f}_{\SM}$ to that of the Laplacian smoothing estimator $\wh{f}_{\LS}$. Using the definition~\eqref{eqn:laplace_operator}, repeated integrating by parts, and then invoking the aforementioned convergence properties of $\Lap_{n,r}$, it can be shown that for any $f \in C^{2s}(\Xset)$ (see \citet{zhou11} for an argument in this spirit),
\begin{equation}
\label{eqn:pointwise_seminorm_convergence}
\abs{f}_{H^s(\Xset)} = \int_{\Xset} \bigl(\Delta_P^sf\bigr)(x) \cdot f(x) \,dx = \lim_{n \to \infty} \frac{1}{n^{s + 1}r^{s(d + 2)}}f^T \Lap_{n,r}^s f.
\end{equation}
This does not imply a formal statement about the convergence of $\wh{f}_{\LS}$ to $\wh{f}_{\SM}$ along the lines of~\eqref{eqn:laplacian_eigenmaps_consistency}. Among other issues, the convergence in~\eqref{eqn:pointwise_seminorm_convergence} is over $f \in C^{2s}(\Xset)$ while the domain of minimization in~\eqref{eqn:smoothing_spline} is $H^{s}(\Xset)$. Nevertheless, it establishes a general connection between $\wh{f}_{\LS}$ and $\wh{f}_{\SM}$.

\subsection{Minimax rates over Sobolev classes}
\label{subsec:minimax_rates}
One reason all of this is intriguing is that, as mentioned previously, the continuum estimators $\wh{f}_{\OS}$ and $\wh{f}_{\SM}$---and test statistics based on functionals of these estimators--- have strong statistical properties. In particular, these estimators and tests are known to have error rates within a constant factor of minimax optimal. Throughout this section, we treat $M > 0$ as a fixed number (i.e. not varying with $n$.)
